---
title: "Reproducible Pitch Presentation Shiny Application"
author: "Genevieve Beart"
date: "15/01/2021"
output: ioslides_presentation

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Summary: COVID-19 Data Visualiser {.smaller}

The purpose of the app developed for this project is to allow users to visualise cumulative COVID-19 data to 13/01/2021. This app would be used for informational purposes as a dashboard to help the user understand the state of the current pandemic. 

The Final Shiny Application is found here: https://circulaic.shinyapps.io/DataProductsFinalAssignment/

The Github Repository with ui.R, server.R and data is here: https://github.com/circulaic/DataProductsFinalAssignment 

This app contains all of the key features required for this assignment: 

* *User input:* two drop-down selection boxes for Continent and Metric
* *Operation on the ui input in server.R:* inputs of Continent and Metric are used in server.R to subset the data for the selected continent, select the correct column to plot given the chosen metric and create the plot
* *Reactive output displayed as a result of server calculations:* the plot updates based on these server calculation
* *Included documentation:* a simple text-based user guide is embedded in the app

##  COVID-19 Data for the App {.smaller}

Data for the app is from the open source Our World in Data Coronavirus Source Data https://ourworldindata.org/coronavirus-source-data. As can be seen below, this data contains day-to-day data on the pandemic for each country. 

``` {r,echo=FALSE, include=FALSE}
library(plotly)
library(lubridate)
library(knitr)
```

``` {r,echo=TRUE, comment=""}
cvd <- read.csv("owid-covid-data-subset.csv", header = TRUE,sep = ",")
cvd$date<-dmy(cvd$date) ##covert dates to date format
head(cvd)
```

##  COVID-19 App Server Calculation {.smaller}

The following server calculations were used in the app: 

``` {r, eval=FALSE, echo=TRUE}
function(input, output) { 
    
    # Fill in the spot we created for a plot 
    output$graph <- renderPlotly({
        continent<-input$Continent  ##brings in the Continent input from the UI
        cvdsub<-cvd[cvd$continent==continent,] ##subsets the data for only the 
              ##selected continent
        
        metric<-input$Metric ##brings in the Metric input from the UI
        ##Returns the column index number of the selected metric:
        colplot<-ifelse(metric=="Total Cases",5,ifelse(metric=="Total Deaths",6,
            ifelse(metric=="Total Cases per Million Population",7,8)))
        
        ##Plots the graph with the selected metric for the selected continent
        p<-plot_ly(x = ~cvdsub$date, y = ~cvdsub[,colplot], type='scatter',
                   mode='line', color=~cvdsub$location)
        layout(p,title="Cumulative COVID-19 Data to 13/01/2021 for Selected Metric
               and Continent", xaxis=list(title="Date"),yaxis=list(title=metric))

    })
}
```

##  COVID-19 App Example Output {.smaller}
The final result in the app is a plot of cumulative COVID-19 figures for a selected metric and continent, broken down by country. For example, if "South America" and "Total Cases per Million Population" were selected, the following plot would appear: - 

``` {r,echo=FALSE, fig_width=4, fig_height=4, message=FALSE, warning=FALSE}
    # Fill in the spot we created for a plot 
        continent<-"South America"
        cvdsub<-cvd[cvd$continent==continent,] ##subsets the data for only the selected continent
        
        metric<-"Total Cases per Million Population" ##brings in the Metric input from the UI
        ##Returns the column index number of hte selected metric
        colplot<-ifelse(metric=="Total Cases",5,ifelse(metric=="Total Deaths",6,ifelse(metric=="Total Cases per Million Population",7,8)))
        
        ##Plots the graph with the selected metric for the selected continent
        p<-plot_ly(x = ~cvdsub$date, y = ~cvdsub[,colplot], type='scatter',mode='line', color=~cvdsub$location)
        layout(p,title="Cumulative COVID-19 Data to 13/01/2021 for Selected Metric and Continent", xaxis=list(title="Date"),yaxis=list(title=metric))



```

